---
output:
  html_document: default
  pdf_document: default
---
title: "Task 3"
author: "Kush Ise"
date: "19/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}

library(tidyverse) 
library(caret)
library(grid)
library(gridExtra)
library(GGally)
library(rpart)
```


```{r}
data
head(data)


```
Data exploration and vizualisation
Distribution of target values
First, let's see how the target values are distributed.

```{r}
iris %>%
    group_by(Species) %>%
    summarise(count = n()) %>%
    ggplot(aes(x = Species, y = count, fill = Species)) + 
        geom_bar(stat = "identity") +
        geom_text(aes(label = as.character(count)), 
                 position = position_dodge(width = 1),
                 hjust = 0.5, vjust = -0.25, size = 3) +
        theme(legend.position = "none") +
        labs(title = "Species", x = "", caption = "Fig.1")
```
Distribution of predictors values

```{r}
h1 <- iris %>%
    group_by(Species, Sepal.Length) %>%
    ggplot(aes(x = Sepal.Length, fill = Species)) + 
        geom_histogram(binwidth = 0.5) + # 
        theme(legend.position = "none") +
        labs(title = "Sepal.Length", x = "", y = "")

h2 <- iris %>%
    group_by(Species, Sepal.Width) %>%
    ggplot(aes(x = Sepal.Width, fill = Species)) + 
        geom_histogram(binwidth = 0.25) + # 
        theme(legend.position = "none") +
        labs(title = "Sepal.Width", x = "", y = "")

h3 <- iris %>%
    group_by(Species, Petal.Length) %>%
    ggplot(aes(x = Petal.Length, fill = Species)) + 
        geom_histogram(binwidth = 1) + # 
        theme(legend.position = "none") +
        labs(title = "Petal.Length", x = "", y = "")

h4 <- iris %>%
    group_by(Species, Petal.Width) %>%
    ggplot(aes(x = Petal.Width, fill = Species)) + 
        geom_histogram(binwidth = 0.5) + # 
        theme(legend.position = "none") +
        labs(title = "Petal.Width", x = "", y = "", caption = "Fig.2")

grid.arrange(h1, h2, h3, h4,
             ncol = 4)
```
```{r}
b1 <- iris %>%
    group_by(Species, Sepal.Length) %>%
    ggplot(aes(x = Species, y = Sepal.Length)) + 
        geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, notch = FALSE) +  
        theme_minimal() +
        theme(legend.position = "none",
             axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0)) +
        labs(title = "Sepal.Length", x = "", y = "")

b2 <- iris %>%
    group_by(Species, Sepal.Width) %>%
    ggplot(aes(x = Species, y = Sepal.Width)) + 
        geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, notch = FALSE) +  
        theme_minimal() +
        theme(legend.position = "none",
             axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0)) +
        labs(title = "Sepal.Width", x = "", y = "")

b3 <- iris %>%
    group_by(Species, Petal.Length) %>%
    ggplot(aes(x = Species, y = Petal.Length)) + 
        geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, notch = FALSE) +  
        theme_minimal() +
        theme(legend.position = "none",
             axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0)) +
        labs(title = "Petal.Length", x = "", y = "")

b4 <- iris %>%
    group_by(Species, Petal.Width) %>%
    ggplot(aes(x = Species, y = Petal.Width)) + 
        geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, notch = FALSE) +  
        theme_minimal() +
        theme(legend.position = "none",
             axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0)) +
        labs(title = "Petal.Width", x = "", y = "", caption = "Fig.3")

grid.arrange(b1, b2, b3, b4,
             ncol = 4)

```


```{r}
f1 <- iris %>%
    group_by(Species, Sepal.Length, Sepal.Width) %>%
    ggplot(aes(Sepal.Length, Sepal.Width, color = Species)) + 
        geom_point() +  
        theme_minimal() +
        theme(legend.position = "none") +
        labs(title = "")

f2 <- iris %>%
    group_by(Species, Petal.Length, Petal.Width) %>%
    ggplot(aes(Petal.Length, Petal.Width, color = Species)) + 
        geom_point() +  
        theme_minimal() +
        theme(legend.position = "none") +
        labs(title = "")

f3 <- iris %>%
    group_by(Species, Petal.Length, Sepal.Width) %>%
    ggplot(aes(Petal.Length, Sepal.Width, color = Species)) + 
        geom_point() +  
        theme_minimal() +
        theme(legend.position = "none") +
        labs(title = "")

f4 <- iris %>%
    group_by(Species, Sepal.Length, Petal.Width) %>%
    ggplot(aes(Sepal.Length, Petal.Width, color = Species)) + 
        geom_point() +  
        theme_minimal() +
        theme(legend.position = "none") +
        labs(title = "", caption = "Fig.4")

grid.arrange(f1, f2, f3, f4, ncol = 2)
```
```{r}
## convert all values to numeric (especially the target)
all_num <- iris
all_num$Species <- as.numeric(all_num$Species)
all_num %>% ggcorr(palette = "RdBu", label = TRUE)
```
Classification using a simple decision tree
Now that we made all those conclusions, let's see how we can determine the Species of an Iris flower using a very simple decision tree.

Using the correlation plot and the top right plot in Fig. 4, we can deduce a set of simple rules. For example:

If Petal.Length < 2.5, Species = setosa

If Petal.Length >= 2.5,

If Petal.Width < 1.8, Species = versicolor

If Petal.Width >= 1.8, Species = virginica
```{r}
mytree <- rpart(Species ~ ., data = iris)
rattle::fancyRpartPlot(mytree, type = 5, sub = "")
```
```{r}
set.seed(3) # for reproducibility
idx <- createDataPartition(factor(iris$Species), 
                           times = 1, # one time
                           p = 0.5, # 50% proportion
                           list = FALSE)

## train set
trainset <- iris[idx,] 
## validation set
validation_set <- iris[-idx,] %>% select(-Species)   
validation_labels <- iris[-idx,] %>% select(Species)
```

```{r}
control <- trainControl(method = "cv", # cross-validation
                        verboseIter = TRUE,
                        savePredictions = "final",
                        classProbs = TRUE) 

set.seed(1)
model <- train(Species ~ ., data = iris,
               trControl = control,
               maximize = TRUE,
               tuneGrid = data.frame(mtry = 4),
               method = "rf")
```
```{r}
model
```
ACCURACY OF PREDICTION MODEL IS 95.33% which is not bad

```{r}
confusionMatrix.train(model)
```
```{r}
val_preds <- predict(model, validation_set)
confusionMatrix(val_preds, validation_labels$Species)
```
```{r}
set.seed(3)
n <- sample(1:nrow(validation_set), size = 3)
validation_set[n,]
```

```{r}
val_preds[n]
```
```{r}
varImp(model)
```

